---
title: "Poisson Class"
author: "Mohd Azmi"
date: "22/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library

```{r}
library(pacman)
p_load(epiDisplay, car, readr, tidyverse, summarytools, broom, ggplot2)
```

# Basic

## x. 2.1-2.3

```{r}
dpois(10, 17.2)  # P(X = x)
ppois(10, 17.2)  # cumulative P(X <= x)
1 - ppois(10, 17.2)  # P(X > x) = 1 - P(X <= x)
```

## plot

```{r}
x <- 0:50
x <- as.data.frame(x)
p.x <- apply(x, 2, function(x) dpois(x, 17.2))
plotA <- cbind(x = x, px = p.x)
plot(p.x, type = "o")
```

```{r}
x2 <- 0:50
x2 <- as.data.frame(x2)
plotA <- x2 %>% mutate(px2 = dpois(x2, 17.2))
plotA
plot(plotA, type="h")
ggplot(plotA, aes(x2, px2)) +
  geom_point() + geom_line() +
  theme_bw()
```




# Analysis

```{r}
mydata <- read_csv("UKaccident.csv")
mydata <- mydata %>% mutate(law2 = if_else(law >= 1, "after", "before"),
                            law2 = as_factor(law2))
mydata
```

```{r}
tapply(mydata$driverskilled, mydata$law2, sum)

table(mydata$law2)
mydata %>% group_by(law2) %>% count()
```

## GLM

```{r}
glm_mod0 <- glm(driverskilled ~ law2, data = mydata, family = poisson)
summary(glm_mod0)
tidy(glm_mod0, conf.int = T)
```

## Model fitness

```{r}
poisgof(glm_mod0)
```

### Standardized Residual

```{r}
pred_glmmod0 <- augment(glm_mod0)
pred_glmmod0[abs(pred_glmmod0$.std.resid) > 1.96,]
```


## Rate Ratio

```{r}
idr.display(glm_mod0)
```

# Numerical Variables

```{r}
myds2 <- read_csv("poisson_sim.csv") %>%
  mutate(prog1 = factor(prog, levels = 1:3, labels = c("General", "Academic", "Vocational")))
myds2
```

```{r}
ggplot(myds2) +
  geom_histogram(aes(num_awards))

ggplot(myds2, aes(x=math)) +
  geom_histogram() +
  facet_wrap(~num_awards)

ggplot(myds2, aes(x=math, y=num_awards)) +
  geom_point()
```

```{r}
myds2 %>% freq(num_awards)
```



## poisson regression model

```{r}
pois_mod0 <- glm(num_awards ~ math, data = myds2, family = poisson)
summary(pois_mod0)
tidy(pois_mod0, conf.int=T)
```

## GOF

```{r}
poisgof(pois_mod0)
```

## standardized residual


```{r}
pred_poismod0 <- augment(pois_mod0)

outlier_predpoismod0 <- pred_poismod0[abs(pred_poismod0$.std.resid) > 1.96,] %>% 
  mutate(pred = exp(.fitted)) %>% select(num_awards, math, pred, .std.resid)
outlier_predpoismod0
```

## Rate Ratio

```{r}
idr.display(pois_mod0)
exp(coef(pois_mod0))
exp(confint(pois_mod0))
tidy(pois_mod0, conf.int = T, exponentiate = T)
```

```{r}
pois_mod0null <- glm(num_awards ~ 1, data = myds2, family = poisson)
anova(pois_mod0null, pois_mod0, test = "Chisq")
```

# Rate Data

## Dataset

```{r}
cigar.day = c(0, 5.2, 11.2, 15.9, 20.4, 27.4, 40.8)
person.yrs = c(1421, 927, 988, 849, 1567, 1409, 556)
cases = c(0, 0, 2, 2, 9, 10, 7)
cig = data.frame(cigar.day, person.yrs, cases)
cig

cig <- cig %>% mutate(rate = cases/person.yrs)
cig
```

## model

```{r}
cig_mod0 <- glm(cases ~ cigar.day, offset = log(person.yrs), data = cig, family = poisson)
summary(cig_mod0)
```

```{r}
pred_cigmod0 <- augment(cig_mod0)
pred_cigmod0
pred_cigmod0a <- pred_cigmod0 %>%
  mutate(pred = exp(.fitted)/person.yrs,
         rate = cases/person.yrs) %>% select(cigar.day, cases, rate, pred)
pred_cigmod0a
```

# Poisson Regression Model

## Dataset

```{r}
myds2
```

## Univariable


### IV math

```{r}
spois_moda <- glm(num_awards ~ math, data = myds2, family = poisson)
summary(spois_moda)
```

### IV Prog

```{r}
spois_modb <- glm(num_awards ~ prog1, data = myds2, family = poisson)
summary(spois_modb)
```

## Update Group

```{r}
myds2 <- myds2 %>%
  mutate(prog2 = fct_relevel(prog1, "Academic"),
         prog2 = fct_collapse(prog2, 'General & Vocational' = c('General', 'Vocational')))

myds2
```

### IV Prog2

```{r}
levels(myds2$prog2)
spois_modc <- glm(num_awards ~ prog2, data = myds2, family = poisson)
summary(spois_modc)
```

```{r}
with(myds2, ctable(num_awards, prog2, prop = "n"))
```

## Multivariable

```{r}
mpois_mod0 <- glm(num_awards ~ math + prog2, data = myds2, family = poisson)
summary(mpois_mod0)
```

### Goodness of fitness

```{r}
poisgof(mpois_mod0)
```

### Rate Ratio or IDR

```{r}
idr.display(mpois_mod0)
tidy(mpois_mod0, exponentiate = T, conf.int = T)
```

### outlier - standardized residual

```{r}
pred_mpoismod0 <- augment(mpois_mod0, type.predict = "class")
pred_mpoismod0[abs(pred_mpoismod0$.std.resid) > 1.96,]
pred_mpoismod0_out <- pred_mpoismod0[abs(pred_mpoismod0$.std.resid) > 1.96,] %>%
  select(math, prog2, num_awards, .fitted, .std.resid)
pred_mpoismod0_out
```

### scaled pearson chi-square

```{r}
quasi_mpoismod0 <- summary(glm(num_awards ~ math + prog2, data = myds2, family = quasipoisson))
quasi_mpoismod0$dispersion
```

### Interaction

```{r}
mpois_mod0_i <- glm(num_awards ~ math + prog2 + math:prog2, data = myds2, family = poisson)
summary(mpois_mod0_i)
```

```{r}
AIC(mpois_mod0, mpois_mod0_i)
anova(mpois_mod0, mpois_mod0_i, test = "Chisq")
```

## Final Model

```{r}
tidy(mpois_mod0, exponentiate = T, conf.int = T)
```

math score increse by 10

```{r}
exp(cbind("Math RR" = coef(mpois_mod0)[[2]]*10,
          "95% LL" = confint(mpois_mod0)[[2]]*10,
          "95% UL" = confint(mpois_mod0)[[5]]*10))
```

